Separation of an upwelling current bounding the Juan de Fuca Eddy, off Vancouver Island

### 2022-02-05 PW13Paper.tex
DOC: rough version to coauthors

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PREAMBLE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% The following two commands will generate a PDF that follows all the requirements for submission
% and peer review.  Uncomment these commands to generate this output (and comment out the two lines below.)
%
% DOUBLE SPACE VERSION FOR SUBMISSION TO THE AMS
\documentclass[11pt]{article}
\usepackage{ametsoc}
% \usepackage{ametsoc2col}
%\linenumbers
\graphicspath{{/Users/jklymak/PW13/figs/mvp/}}
\usepackage[plain]{fancyref}
\usepackage{xcolor}
\usepackage{lineno}
\newcommand{\tempS}[1]{}
\newcommand{\twowidth}[0]{6.4in}
\newcommand{\onewidth}[0]{3.2in}
\newcommand{\mn}[1]{{\sc #1}}
\DeclareGraphicsExtensions{%
    .pdf,.png}
%
% The following two commands will generate a single space, double column paper that closely
% matches an AMS journal page.  Uncomment these commands to generate this output (and comment
% out the two lines above. FOR AUTHOR USE ONLY. PAPERS SUBMITTED IN THIS FORMAT WILL BE RETURNED
% TO THE AUTHOR for submission with the correct formatting.
%
% TWO COLUMN JOURNAL PAGE LAYOUT FOR AUTHOR USE ONLY
%%%%\documentclass[10pt]{article}
%%%%\usepackage{ametsoc2col}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%
% Enter your Abstract here
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\linenumbers
\newcommand{\myabstract}{}
%
\begin{document}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE
%
% Enter your TITLE here
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{\large{Evidence for cross-shelf exchange driven by a submarine canyon}}}
%
% Author names, with corresponding author information.
% [Update and move the \thanks{...} block as appropriate.]
%
\author{\textsc{Jody M. Klymak,}
				\thanks{\textit{Corresponding author address:}
				Jody Klymak, University of Victoria, Victoria, BC, Canada
				\newline{E-mail: jklymak@uvic.ca}}\quad
				\\
\textit{\footnotesize{School of Earth and Ocean Sciences and Department of Physics and Astronomy, University of Victoria, Victoria, Canada}}
}
%
% Formatting done here...Authors should skip over this.  See above for abstract.
\ifthenelse{\boolean{dc}}
{
\twocolumn[
\begin{@twocolumnfalse}
\amstitle

% Start Abstract (Enter your Abstract above.  Do not enter any text here)
\begin{center}
\begin{minipage}{13.0cm}
\begin{abstract}
	\myabstract
	\newline
	\begin{center}
		\rule{38mm}{0.2mm}
	\end{center}
\end{abstract}
\end{minipage}
\end{center}
\end{@twocolumnfalse}
]
}
{
\amstitle
\begin{abstract}
\myabstract
\end{abstract}
\newpage
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN BODY OF PAPER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Cross-shelf exchange is important to the health and productivity of continental shelf regions, where offshore nutrients and oxygenated water is exchanged with nutrient and oxygen depleted near-shore water.  Cross-shelf transport requires a level of ageostrophic flow since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange.  Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.  Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface

%\citet{klymaketal10a}

@@

Data collection was between 21 and 30 August, 2013, aboard the \emph{R/V Falkor}.  The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.

@@

     \includegraphics[width=6.2in]{LocMapBoth.pdf}
    \caption{a) Study site on the Vancouver Island Shelf.  The blue arrow indicates the direction of the Vancouver Island Current.  The pink arrow indicates southward flow of the coastal upwelling current.  The dashed box indicates the approximate limits of the study area.  b) The study area with standard hydrographic casts indicated in yellow dots, and Moving Vessel Profiler casts as red dots.
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProc.ipynb ;
        /Users/jklymak/PW13/figs/mvp//LocMapBoth.pdf}
      \label{fig:LocMapBoth} }
  \end{center}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{SpiceO2264}
    \caption{Spatial overview of a) local spice (temperature anomaly on an isopycnal: see text), and b) oxygen saturation on the $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.  Grey cross-slope lines are cross sections indicated in \fref{fig:CrossSectionsSpice}.  Along-slope grey lines are every 20 km in the cross-slope direction, with $x=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.

      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
        /Users/jklymak/PW13/figs/mvp//SpiceO2264.pdf}
      \label{fig:SpiceO2264} }
  \end{center}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{CrossSectionsSpice}
    \caption{
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew.ipynb ;
        /Users/jklymak/PW13/figs/mvp//CrossSectionsSpice.pdf}
      \label{fig:CrossSectionsSpice} }
  \end{center}
\end{figure*}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=3in]{TsLineDefn}
    \caption{
      \tempS{\footnotesize /Users/jklymak/PW13/figs/mvp///MvpProcNew ;
        /Users/jklymak/PW13/figs/mvp//TsLineDefn.pdf}
      \label{fig:TsLineDefn} }
  \end{center}
\end{figure}

%\begin{appendix}

%\subsection{First appendix secondary heading}
%
%\subsection{Second appendix secondary heading}
%
%\subsubsection{First appendix tertiary heading}
%
%\subsubsection{Second appendix tertiary heading}
%
%\paragraph{First appendix quaternary heading}
%
%\paragraph{Second appendix quaternary heading}
%
%\end{appendix}

% Create a bibliography directory and place your .bib file there.
% -REMOVE ALL DIRECTORY PATHS TO REFERENCE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
\ifthenelse{\boolean{dc}}
{}
{\clearpage}
\bibliographystyle{ametsocjmk}
\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FIGURES-REMOVE ALL DIRECTORY PATHS TO FIGURE FILES BEFORE SUBMITTING TO THE AMS FOR PEER REVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
% \graphicspath{{/Users/jklymak/Dropbox/PW13/satellite/doc/},
%               {/Users/jklymak/Dropbox/PW13/figs/mvp/},
%               {/Users/jklymak/Dropbox/PW13/analysis/mvp/doc/},
%               {/Users/jklymak/Dropbox/PW13/LaPerouse13/doc/},
%               {/Users/jklymak/Dropbox/PW13/auxdata/doc/}}

\graphicspath{{./figs/}}
\DeclareGraphicsExtensions{%
              .pdf,.png}

\usepackage[plain]{fancyref}
% \usepackage[]{hyperref}
%\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

\draftfalse

\journalname{JGR: Oceans}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Separation of an upwelling current bounding the Juan de Fuca Eddy, off Vancouver Island}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{Jody M. Klymak\affil{1} and Everyone }

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}
\affiliation{1}{School of Earth and Ocean Sciences and Department of Physcics \& Astronomy, University of Victoria, B.C.\ Canada}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

\correspondingauthor{J.\ Klymak}{jklymak@uvic.ca}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item The shelfbreak current along Vancouver Island separates downstream of a submarine bank.
\item Offshore water is drawn onto the shelf and forms a sharp semi-persisnte front with the Juan de Fuca eddy.
\item The Juan de Fuca eddy shows evidence of long residence times, and little evidence of deep-water origin.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}

\end{abstract}

\section*{Plain Language Summary}

The southern Vancouver Island continental shelf is very productive due to high nutrient input from the Strait of Juan de Fuca and Salish Sea estuarine system, and substantial cross-shelf transport due to the complicated topography.  The estuarine water flows poleward, hugging the coast as the Vancouver Island current, and in the summer shelf water flows equatorward.  Trapped between these two currents is the Juan de Fuca eddy, a stagnant region of water that occupies the shelf where it abruptly widens just north of the Strait of Juan de Fuca.

Here we present intensive sampling of the Juan de Fuca eddy region. The observations show the eddy is low in oxygen, and we find has undergone substantial vertical and lateral mixing. This indicates that the waters have a surprisingly long residence time for a coastal feature, and that this region extends the depth of the water column.  In contrast to previous literature we find that the water likely has low oxygen from being used for respiration rather than being pulled from deep in the California Undercurrent.

The sampling also shows a remarkable flow separation of the equatorward shelf current  The current is shown to detach and is pushed offshore. Such events are readily seen in satellite imagery, but the results here indicate that the separation extends the depth of the water column on the shelf, and that this separation may be partially driven by the local bathymetry.  The separation is a very strong cross-shelf exchange event, and likely transports substantial nutrient-rich coastal water offshore to drive productivity in the deeper ocean adjacent to the continental slope.

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MAIN BODY OF PAPER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Cross-shelf exchange is important to the health and productivity of continental shelf regions, where oxygenated water is exchanged with nutrient-rich, but oxygen-depleted nearshore water.  Cross-shelf transport usually requires ageostrophic flow, since balanced flow will tend to follow topographic contours, often providing a barrier to lateral exchange \cite{brink16}.  Mechanisms for cross-shelf exchange include internal waves and instabilities in shelf-break fronts.  Most-dramatically, however, topography can drive cross-slope exchanges, often seen most clearly in surface imagery, but occasionally observed \emph{in-situ} \cite{barthetal00}.

Here we present detailed observations from the southern Vancouver Island shelf, collected in summer 2013, and sampled with a rapid profiling vehicle equipped with a CTD and oxygen sensor, supplemented by traditional hydrographic surveys. The shelf has complicated bathymetry \fref{fig:LocMapBoth}, with a somewhat typical shelf poleward of our study site, but then widening in the lee of La Perouse Bank, and then hitting the Juan de Fuca canyon, equatorward.  Water from the strait of Juan de Fuca flows poleward as a buoyant current that hugs the coast \cite{thomsonetal89, hickeyetal91}, while shelf water flows from poleward in the summer under upwelling conditions, both forced by local winds and via tele-connections with the long, homogenous shelves equatorward off Washington and Oregon \cite{hickeyetal91,thomsonkrassovski15,engidaetal16}.  Trapped between these two currents is a region of relatively homogenous water that has been termed the Juan de Fuca eddy \cite{freelanddenman82,freelandmcintosh89,foremanetal07,macfadyenhickey10}.

One of our findings below is that the water in the Juan de Fuca eddy is apparently quite old.  The low oxygen found in the Juan de Fuca eddy has caused it to be inferred to be upwelled from as deep as 400 m \cite{freelanddenman82,deweycrawford88}.  This inference was based on the oxygen being conservative, and hence having to come from close to the oxygen minimum zone found further offshore \cite{mackasetal87}.  This lead to hypotheses that perhaps the eddy was fed by waters being drawn up a spur canyon of the Juan de Fuca canyon (often called the Tully Canyon) via ageostrophic transport from offshore to onshore due to the low pressure in the eddy center \cite{weaverhsieh87}.  Below, we will argue that there is not substantial evidence of such transport, and that oxygen is likely low because of consumption due to respiration.

A second important finding is that the water flowing equatorward along the shelf from upstream is all displaced from the shelf and pushed offshore and replaced by offshore water that floods the southern shelf.  Such events are important as they tend to transport nutrient rich shelf water out into the open ocean, fueling offshore ecosystems.
This cross-shelf exchange has not, to our knowledge, been remarked upon for the Vancouver Island shelf.  Mesoscale separations have been observed further to the south off Cape Blanco \cite{barthetal00}, and was also inferred at other locations along the coast.  Similar observations have been made elsewhere \cite[e.g.]{relvasbarton05}, though usually in the presence of a distinct coastal headland, a feature not found at our study location.  The mechanisms driving such separations are poorly understood, but hypotheses include coastal hydraulics \cite{dalebarth01} leading to along-shore trapping of coastal waves,  wind stress curl variations \cite{castelaobarth07}, and induced relative vorticity due to stretching of parcels that inertially overshoot their initial isobaths due to a sudden change in downstream bathymetry.

In this paper we present our observations from 2013 by describing our at-sea observation tools (\fref{sec:Site}), presenting the data in a number of ways to highlight the important processes, with a focus on the offshore side of the eddy region (\fref{sec:Observations}) where we consider the age of the eddy, the steadiness of the front separating the eddy and the offshore water, properties along the spur canyon that was believed to feed the eddy, and finally describe the large offshore separation of the shelf-break current.  We discuss the origin of the eddy and the implications of the separating jet.

@@

Observations from three summer cruises in 2013 are discussed here. Two were hydrographic surveys carried out on the \emph{CCGS Tully}, one in June, and the other in late September.  The third was a finescale survey made in the Vancouver Island shelf between 21 and 30 August 2013, aboard the \emph{R/V Falkor}.  The study site was the southern portion of the Vancouver Island Shelf (\fref{fig:LocMapBoth}a), a particularly complicated region due to the bathymetry and varied forcing.

\begin{figure*}[htbp]
  \includegraphics[width=5in]{LocMapBoth.pdf}
  \caption{a) Study site on the Vancouver Island Shelf.  The blue arrow indicates the direction of the Vancouver Island Current.  The pink arrow indicates southward flow of the coastal upwelling current.  The dashed box indicates the approximate limits of the study area.  b) The study area with hydrographic casts from the La Perouse cruises along the LB and LC Lines are green dots, hydrographic casts during the Falkor cruise indicated in yellow dots, and Moving Vessel Profiler casts as red dots.  The coordinate system used for this paper is shown with alternating grey and white bands at 10-km intervals.}
  \label{fig:LocMapBoth}
\end{figure*}

At the south end of the study site is the Juan de Fuca canyon, which feeds dense water into Juan de Fuca Strait.  This dense water is mixed with fresh water from the Fraser River at the sills and archipelagos further inland, and fluxes out the Strait again, where it turns poleward along Vancouver Island to form the Vancouver Island Current.  The Juan de Fuca Canyon has a notable spur canyon, the Tully Canyon, that incises the shelf towards the north into the study site (\fref{fig:LocMapBoth}b).  The rest of the shelf is punctuated by a series of banks and shallow basins.  Notable is La Perouse bank, running along the shelf from the north, and moving isobaths offshore until it abruptly ends at around 48.5 N. The shelf break is characterized by a series of submarine canyons, in particular Nitnat Canyon and Barkley Canyon.

Observations from the LB and LC lines were collected aboard the \emph{CCGS Tully} from 2013-05-30 to 2013-05-31, which we will call ``June'', and 2013-09-07 to 2013-09-09, which we will call ``September''.  These lines span the depths from about 50 m to deep offshore.  The data comes from a lowered Seabird 9-11 CTD, with an SBE 43 oxygen sensor.  Oxygen data have been checked against bottle casts, and the CTD corrected for sensor offsets and thermal lags.

Observations aboard the \emph{R/V Falkor} were carried out from 2013-08-21 to 2013-08-30, starting two weeks before the September Tully cruise.  Data were collected with an ODIM Brooke-Ocean Moving Vessel Profiler, profiling to 200 m.  The MVP was equipped with an AML Oceanographic CTD, and a Rinko Oxygen sensor with a 7-s response time foil.  The MVP profiles at vertical fall speeds of $3\ \mathrm{m\,s^{-1}}$ to depths of 200 m or to within 5 m of the seafloor, whichever is shallower.  Data collection took place while the ship cruised at speeds between 5 and 8 kts, usually at around 6 kts to enable fine horizontal spacing of the casts, with typical spacing less than 800 m.  Data is collected during down- and up-casts, but is only reported for down-casts because the path through the water is highly asymmetric on the up-cast, whereas the downcast is largely vertical.  The rapid speed of the profiling makes the oxygen measurements somewhat coarse, and probably biased due to the phase lag of the sensor, so we treat these largely qualitatively in this paper.

Unfortunately, neither vessel had an operational acoustic Doppler profiler during the cruises with which to make current measurements.

Winds during the cruises were typical for the west coast of Vancouver Island, with down-welling favorable winds during July and early August, and then intermittent winds before and during the Falkor cruise.  Note that this locale is strongly affected by coastally trapped waves from further south, so upwelling usually persists despite local wind forcing \cite{thomsonkrassovski15, engidaetal16}.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=5.5in]{LaPeWind}
    \caption{
      Wind from the La Perouse buoy \cite{DFOWind2013C46206}.  The vertical direction is along-shelf (chosen as a heading of 315 N).  The wind components has been low-pass filtered to one-day averages.  The timing of the cruises discussed in the paper are shown as colored bands.
      \label{fig:LaPeWind} }
  \end{center}
\end{figure*}

\section{Observations}
\label{sec:Observations}

\subsection{Early and late summer broad-scale surveys}

Hydrographic sections along the LB and LC lines highlight summer conditions on the Southern Vancouver Island Shelf, and indicate some of the features we are focusing on in this paper (\fref{fig:LaPerouse2013Ctd}).  During both cruises and at both lines there is clear evidence of upwelling, with the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal reaching from 130 m to shallower than 85 m over the shelf.  This dense water tends to be low in oxygen and cool. Further onshore, isopycnals tilt back down, and warm water fills the water column, evidence of the Vancouver Island Current that hugs the coast here.

\begin{figure*}[htbp]
  \begin{center}
     \includegraphics[width=130mm]{LaPerouse2013Ctd}
     \caption{Observations along LC and LB lines in Jun and Sep 2013.  X is across-shelf distance in the coordinate system shown in \fref{fig:LocMapBoth}. Potential density is contoured every $0.2\ \mathrm{kg\,m^{-3}}$, with the $26.4\ \mathrm{kg\,m^{-3}}$.  Potential temperature $\theta$, spice anomaly $\gamma$ and oxygen concentration $O_2$ are colored for each section.}
     \label{fig:LaPerouse2013Ctd}
  \end{center}
\end{figure*}

The ``Juan de Fuca Eddy'' region is along the LB Line, mostly inshore of 0 km. This water is cooler and lower in oxygen than along the same isopycnal further offshore.  Also note that during the June cruise the water in this region was cooler and more oxygenated than later in the summer during the September cruise.  Near the surface, the isopycnals are domed, but the low-oxygen anomaly extends as high as the thin surface mixed layer.

In contrast, the water upstream, along the LC line, does not show as much drawdown of oxygen, and is somewhat warmer, at least offshore of La Perouse bank (\fref{fig:LaPerouse2013Ctd} top three rows).  Water does not appear to make it over La Perouse bank into the basin onshore, or if it does, it does so intermittently.  There appears to be some mixing of the deeper water and shallow water near $26.4\ \mathrm{kg\,m^{-3}}$, with the water along this isopycnal becoming cooler and lower in oxygen where it intersects the shelf.

The $\theta$--$S$ properties show many of the features demonstrated above (\fref{fig:LaPerouse2013TS}a).  Below $26.6\ \mathrm{kg\,m^{-3}}$ or so, the deep water masses are found offshore, and are largely homogenous. Shallower, there are distinct differences between the water found on the shelf and water found more offshore.  These anomalies are hard to parse directly, so in this paper we use a spice anomaly, relative to a straight line in $\theta$--$S$ space (\fref{fig:LaPerouse2013TS}). The line passes the points $7.75\ \mathrm{^oC}$, $30\ \mathrm{psu}$ and $6.6\ \mathrm{^oC}$, $35\ \mathrm{psu}$, and spice anomaly for a given water sample at density $\sigma_{\theta}$ is $\gamma = \alpha \left(T - T_0\left(\sigma_{\theta}\right)\right) + \beta \left(S - S_0\left(\sigma_{\theta}\right)\right)$ where $T_0\left(\sigma_{\theta}\right)$ and $S_0\left(\sigma_{\theta}\right)$ are the temperature and salinity along the mixing line at the same density as the water sample, with the sign convention that positive is warmer and saltier than data along the mixing line.  It is hard to see from the $\theta$--$S$ plot in \fref{fig:LaPerouse2013TS}, but this mixing line is a strong feature of the water masses discussed in this paper.

\begin{figure}[htbp]
  \begin{center}
     \includegraphics[width=5in]{LaPerouse2013TS}
     \caption{Potential temperature versus salinity for the La Perouse data shown in \fref{fig:LaPerouse2013Ctd}.  Potential density is contoured in the same way.  A definition of spice anomaly is shown in this plot, and discussed in the text.  The rectangle is the $\theta-S$ range used in \fref{fig:TSdensSpice} below.}
     \label{fig:LaPerouse2013TS}
  \end{center}
\end{figure}

The spice anomaly defined this way is mapped in the middle row of each set of plots in \fref{fig:LaPerouse2013Ctd}.  Offshore water tends to be high is absolute spice anomaly, with a positive-anomaly layer (red) centered at $26.4\ \mathrm{kg\,m^{-3}}$ sandwiched between negative-anomaly layers above and below.  On the shelf ($-10\, \mathrm{km} < X < 30\, \mathrm{km}$), the distinct $\theta$--$S$ masses are attenuated and much closer to the mixing line than the offshore water.  Hugging the coast ($X > 30\, \mathrm{km}$), is the Vancouver Island Current, with a negative spice anomaly in June, and a positive anomaly in summer, consistent with the warming of this water during the summer.

The June and September cruises are quite similar in water properties, particularly in the offshore waters.  Water has upwelled from deeper depths in September, but has most of the same water mass characteristics as earlier in the year.  The water on the shelf is somewhat warmer than earlier in the year, and by September has started to fall off the mixing line.

In terms of oxygen, both cruises have a similar dichotomy between the onshelf and offshelf water, with water found in the eddy tracking at least $100\ \mathrm{\mu mol\, kg^{-1}}$ lower on the shelf than offshelf (\fref{fig:LaPerouse2013Ctd}, \fref{fig:LaPerouse2013TS}b).  The very deepest water ($S\approx 33.9\ \mathrm{psu}$) on the shelf shows a further $50\ \mathrm{\mu mol\, kg^{-1}}$ drop between June and September along the LB line.

\subsection{Fine-scale surveys}

\subsubsection{Overview}

As shown in \fref{fig:LocMapBoth},  we conducted a large spatial survey of the shelf using the Moving Vessel Profiler, collecting over 2000 CTD casts in a relatively synoptic manner.  This amount of data allows a reasonable census of the water masses that motivated our choice of the mixing line shown in \fref{fig:LaPerouse2013TS}.  Binning in temperature and salinity, we can look at sample frequency of these observations \fref{fig:TSdensSpice}, and we see that there are distinct water masses with few observations between them. Looking along the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal, we see that there is a large population along the mixing line.  At the opposite extreme, we see that there is a large population with a positive spice anomaly of approximately $0.2\ \mathrm{kg\,m^{-3}}$, representing the offshore water.  Between these two end points there is a third water mass, somewhat indistinct from the offshore water mass, but very distinct from the water mass along the mixing line.  These three water masses exist on deeper and shallower isopycnals as well, but become hard to distinguish from $\theta$--$S$ properties alone as the properties converge near $26.1\,\mathrm{kg\,m^{-3}}$.

% TODO: shallower?

@@

     \includegraphics[width=3.5in]{TSdensSpice.pdf}
    \caption{Sample density of salinity and potential density data from the cruise, with a logarithmic color scale.  Grey contours are potential density relative to the surface, every $0.1\ \mathrm{kg\,m^{-3}}$, and the magenta contour is the $26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.  Spice anomaly, $\gamma$ is defined as temperature anomaly scaled by $\alpha$ along isopycnals from the mixing line found in the well-mixed region onshore, in $\mathrm{kg\,m^{-3}}$.
    \label{fig:TSdensSpice}}
  \end{center}
\end{figure*}

These three water masses are quite coherent in space.  While we did not do the specific La Perouse sections, our spatial coverage is such that we can make consistent synthetic sections (\fref{fig:CrossSectionsSpice}), taken at $y=42, 28, 14$, and $0\, \mathrm{km}$ (see \fref{fig:LocMapBoth}).  During the August cruise, the upstream water along $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ has moderate spice anomaly of  $\gamma \approx 0.1\,\mathrm{kg\,m^{-3}}$ (\fref{fig:CrossSectionsSpice}a), and in this representation, the water appears "pink".  There is some higher-anomaly water offshore of $x \approx -40\, \mathrm{km}$ along this upstream section.

By $y\approx 14\, \mathrm{km}$, this ``pink'' water along $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ is gone and is replaced by high spice anomaly water $\gamma \approx 0.2 \,\mathrm{kg\,m^{-3}}$ (``red'') that is directly next to well-mixed water ($\gamma \approx 0.0 \,\mathrm{kg\,m^{-3}}$, ``white''), a situation that persists further down the shelf (\fref{fig:CrossSectionsSpice}c, d).  The onshore waters ($-5 < X < 40\, \mathrm{km}$) are remarkably well-mixed in $\theta$--$S$ space, despite still retaining significant density stratification.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{CrossSectionsSpice}
    \caption{Synthetic sections of spice anomaly from the MVP data, projected along four lines from poleward (upstream of coastal current) a) to equatorward (downstream); the lines are shown in \fref{fig:LocMapBoth}b).    Isopycnals are contoured every $\sigma_{\theta} = 0.2\,\mathrm{kg\,m^{-3}}$ , with the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal highlighted in magenta.  Colors are spice anomaly as defined in \fref{fig:TSdensSpice}. "BC" in panel (c) refers to Barkley Canyon.
      \label{fig:CrossSectionsSpice} }
  \end{center}
\end{figure*}

Oxygen profiles, computed in the same manner, show similar trends, with the low-spice anomaly water clearly having a very low oxygen signature, while the strong negative spice anomaly shallower than $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ has very high oxygen, and the deeper water is also more oxygenated.  Of note is that compared with the coarser sampling along the LC Line (\fref{fig:LaPerouse2013Ctd}), we see that there is a very abrupt transition between the low spice anomaly/low-O2 water and the offshore water over the ``Eddy'' region.  We will see below that this transition is even sharper than indicated by these synthetic sections, since they alias some lateral and temporal irregularity in the $\theta$--$S$-compensated front.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{CrossSectionsO2}
    \caption{Synthetic sections of Oxygen, approximate percent saturation from the MVP data.
      \label{fig:CrossSectionsO2} }
  \end{center}
\end{figure*}

The spatial patterns are very clear when considering a map-view of properties along the $\sigma_{\theta} = 26.4\,\mathrm{kg\,m^{-3}}$ isopycnal (\fref{fig:SpiceO2264}).  There is a region onshore of the south tip of La Perouse Bank (approximately 44.5 N, 125.75 W), that consists of water that is found along the mixing line (spice anomaly $\gamma \approx 0 \mathrm{kg\,m^{-3}}$) with low oxygen saturations; this water is called the Juan de Fuca eddy (or Tully eddy), though here an eddy shape is not readily apparent.  Offshore of this region is water that is very warm and salty in comparison ($\gamma > 0.15 \mathrm{kg\,m^{-3}}$), and relatively high in oxygen (saturations of approximately 60\%).  The region between these two water masses is very abrupt, and is concave with respect to the isobaths on the shelf,  stretching from La Perouse Bank to a shallow bank just above the Juan de Fuca canyon (48.3 N and 125.4 W).

Upstream of La Perouse bank, and Barkley Canyon, the water along the shelf has a weaker spice anomaly than the water further downstream ($\gamma \approx 0.7$, light pink colors).  This water is also somewhat lower in oxygen than water from offshore, though not as depleted as the water in the "eddy".  This water appears to be pushed offshore just upstream of Barkley Canyon, and replaced on the shelf by the warmer offshore water.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=6.2in]{SpiceO2264}
    \caption{Spatial overview of a) local spice (temperature anomaly on an isopycnal: see text), and b) oxygen saturation on the $\sigma_{\theta} = 26.4\ \mathrm{kg\,m^{-3}}$ isopycnal.  Grey cross-slope lines are cross sections indicated in \fref{fig:CrossSectionsSpice}.  Along-slope grey lines are every 20 km in the cross-slope direction, with $x=0\ \mathrm{km}$ near the 100-m isobath at the north end of the observation area.
   \label{fig:SpiceO2264}
    }
  \end{center}
\end{figure*}

\subsubsection{Frontal survey}

A systematic survey through the front between the offshore water and the ``eddy'' water demonstrates the structure of the front (\fref{fig:Frontsurvey}).  The survey started close to shore and passed through the Vancouver Island Current (casts 2030--2080).  This water is buoyant and signficiantly fresher and cooler than water at the same density near the surface.  There is strong evidence of isopycnal mixing of the surface water and the outflow water, with warm-salty water in the relatively wide front between the ``eddy'' water and the water in the current.

The front between the offshore water and the ``eddy'' water was passed through 10 times, showing its evolution following the along-shore equatorward flow.  First, as noted in the composite sections, isopycnals slope up from offshore onto the shelf.  This is in agreement with a shelf-break current, though the isopycnals flatten out in the ``eddy'' region.  The front starts out very sharp, (along-track distance $50\ \mathrm{km}$) though two small tendrils can be seen separating from the front on the inshore side.  Similar tendrils are found on the second pass, perhaps a bit more separated from the front ($\approx 80\ \mathrm{km}$), and on the third pass ($\approx 95\ \mathrm{km}$).  Note that these tendrils are made of up partially mixed water.  The subsequent passes have more of this partially mixed water, such that the partially mixed region is up to 5-km wide by the fifth pass ($\approx 150\ \mathrm{km}$).  However, note that deeper isopycnals retain a sharp front, and indeed the front appears sharp again by the seventh pass at all depths.
